Processing Notebook

### 2023-06-13 manuscript/processing-NB.qmd
Experiments in model fitting

1. Process the raw data output from PAIPR (performed in `scripts/preprocess.py`)
2. Compute topographic data necessary for later analyses (performed in `scripts/topo.calc.R`)
3. Clean the data and re-format for `R` processing (performed in `scripts/Rdata-prep.R`)
4. Build, run, and validate the `INLA` model (`scripts/inla-model.R`)
5.

>>>>

1. Process the raw data output from PAIPR (performed in `scripts/preprocess.py`)
2. Compute topographic data necessary for later analyses (performed in `scripts/topo.calc.R`)
3. Clean the data and re-format for `R` processing (performed in `scripts/Rdata-prep.R`)
4. Build, run, and validate the `INLA` model (`scripts/inla-model.R`)

## Analysis

I first import the requisite libraries

```{r}
library(here)
library(INLA)
library(dplyr)
library(ggplot2)
library(terra)
library(colorspace)
library(sf)
```

and the model results generated previously.

```{r}
# Load necessary data for analysis and plotting (generated using `scripts/article-script.R)
mod.full = readRDS('/media/durbank/WARP/Research/Antarctica/WAIS-central/data/interim-models/mod-full.skip10.cut5.rds')
valid.df = readRDS('/media/durbank/WARP/Research/Antarctica/WAIS-central/data/dfs/valid.df.rds')
results.df = readRDS('/media/durbank/WARP/Research/Antarctica/WAIS-central/data/dfs/results.df.rds')
grid.rast = rast('/media/durbank/WARP/Research/Antarctica/WAIS-central/data/rasters/grid.tif')
mu.rast = rast('/media/durbank/WARP/Research/Antarctica/WAIS-central/data/rasters/mu.tif')
sd.rast = rast('/media/durbank/WARP/Research/Antarctica/WAIS-central/data/rasters/sd.tif')
```

I define a helper function to transform the posteriors to more intuitive scales for improved plots.
```{r}
t.marg = function(mod.marg, t.func, param.names=NULL, n=100) {

  if (any(class(mod.marg) == "matrix")) {
    mod.marg = list(mod.marg)
    if (!is.null(param.names)) {
      names(mod.marg = param.names)
    } else {
      names(mod.marg) = "UNNAMED"
    }
  }

  if (is.null(param.names)) {
    param.names = names(mod.marg)
  }

    if (is.list(t.func) != TRUE) {
    t.func = list(t.func)
  }
  if (length(t.func) == 1) {
    t.func = rep(t.func, length(param.names))
  }

  # Preallocate output tbl
  out = tibble(x=numeric(), y=numeric(), Param=vector(mode="character"))

  marg.names = names(mod.marg)

  for (i in 1:length(marg.names)) {
    marg.i = inla.tmarginal(t.func[[i]], mod.marg[[marg.names[i]]], n=n)
    out = out %>% bind_rows(tibble(x=marg.i[,1], y=marg.i[,2], Param=param.names[i]))
  }
  return(out)
}
```

```{r}
# Extract parameter posteriors
marg.fixed = t.marg(mod.full$marginals.fixed, function(x) exp(x))
HP.names = c("Gamma precision", 
             "Time precision", "Rho for time", 
             "Range for spat.idx", "Stdev for spat.idx")
marg.hyper = t.marg(mod.full$marginals.hyperpar, function(x) (x), param.names = HP.names)
p.selection = c("Intercept", "dem", "Gamma precision", 
                "Time precision", "Rho for time", 
                "Range for spat.idx", "Stdev for spat.idx")
posts = marg.fixed %>% bind_rows(marg.hyper) %>% filter(Param %in% p.selection)
```

```{r}
#| label: fig-ParamPost
#| fig-cap: "Posterior distributions of model parameters. Blah blah blah."
ggplot(posts, aes(x=x,y=y)) + geom_line() + 
  facet_wrap(vars(Param), scales = "free") + 
  theme(text = element_text(size = 12))  
```

```{r}
#| label: fig-trends
#| fig-cap: "Posterior results for accumulation trends across the region."
#| fig-subcap: 
#|   - "Mean trends."
#|   - "Standard deviation of trends"
#| layout-ncol: 2

plot(grid.rast$trend)
plot(grid.rast$trend.sd)
```
